Effects of pulsed hollow electron lens operation on the beam core in HL-LHC: \\First experimental studies and simulations.

### 2018-01-11 resex.tex
remove grid in plots

\section{Hollow electron lens for HL-LHC\label{sec:hel}}
\subsection{General Overview\label{sec:hel:intro}}
In general, electron lenses are DC or pulsed low energy electron beams. The electron beam is generated in an electron gun, then guided and confined by strong solenoids and finally dumped onto a collector. Exemplary for the conceptual design of all electron lenses, the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
\begin{figure}[h]
	\includegraphics[width=1.0\linewidth]{hel_layout_epbeam}% Here is how to import EPS art
	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL.}
\end{figure}
The circulating beam, in case of the LHC the proton beam, is then affected by the electromagnetic field of the electron beam. For the application of active halo control, the electron beam needs to generate an electromagnetic field at the location of the halo particles while leaving the core untouched. This can be achieved by using a uniform hollow distribution in radius $r=\sqrt{x^2+y^2}$ with inner radius $R_1$ and outer radius $R_2$. In this case, the circulating proton beam experiences a radial kick $\theta(r)$
\begin{equation}\label{eq:field_1}
\theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max},
\end{equation}
where $f(r)$ is a shape function with
\begin{equation}\label{eq:field_2}
f(r) =
\begin{cases} 0 &,\quad r< R_1,\\
\frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2,\\
1 &,\quad R_2 \leq r
\end{cases}
\end{equation}
and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
\begin{equation}
\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{q}{p_0}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
\end{equation}
where $L$ is the length of the e-lens, $I_T$ the total electron beam current, $\beta_{e}$ and $\beta_{p}$ the relativistic $\beta$ of electron and proton beam, $\frac{q}{p_0}=\left(B\rho\right)_p$ the magnetic rigidity for the reference particle, $c$ the speed of light and $\epsilon_0$ the vacuum permittivity.
\begin{figure}[b]
	\includegraphics[width=0.8\linewidth]{hel_field}% Here is how to import EPS art
	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
\end{figure}
\begin{table}[t]
	\caption{\label{tab:hel_param}%
		HL-LHC hollow electron lens parameters as in~\cite{hel_cdr}.
	}
	\begin{ruledtabular}
		\begin{tabular}{lcc}
			Geometry & Value& Unit\\
			\colrule
		Length $L$    &  3 & m\\
		Desired range of scraping positions\footnote{$\sigma$ denotes here the Beam sigma and not the collimation $\sigma$ which assumes a normalized emittance of 3.5~$\mu$m.} & 4-8 &$\sigma_p$\\
		\colrule
		Magnetic fields & & \\
		\colrule
		Gun solenoid, $B_g$ & 0.2-0.4 & T\\
		Main solenoid, $B_m$ & 2-6 & T\\
		Collector solenoid, $B_c$ & 0.2-0.4 & T\\
		Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
		\colrule
		Electron gun & & \\
		\colrule
		Peak yield $I_e$ at 10~keV & 5.0 & A\\
		Gun perveance $P$ & 5 & $\mu$perv\\
		Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
		\colrule
		High voltage modulator & & \\
		\colrule
		Cathode-anode voltage & 10.0 & kV\\
		Rise time (10\%-90\%) & 200 & ns \\
		Repetition rate & 35 & kHz
		\end{tabular}
	\end{ruledtabular}
\end{table}

@@

\subsubsection{Prediction of the most efficient pulsing pattern\label{sec:pattern}}
\begin{figure}[b]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_sigm.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_emit1.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2u_pattern_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors. The excitation was applied in both planes with an amplitude of 96~nrad.}
\end{figure}
In order to predict the most efficient pulsing pattern, the loss rate and emittance growth from the Distribution Tracking are compared for the different pulsing patterns. The results for the 2016 optics and tune are shown in Fig.~\ref{fig:patternsim}. The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and the random excitation. As the bunch length decreases accordingly, the losses are mostly associated to off-momentum particles. Only for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as random excitation a considerable emittance growth is visible, where the random excitation has by far the strongest effect compared to any $k^{\mathrm{th}}$~turn pulsing pattern.

@@

\begin{figure}[t]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_7thhv_3_5um_hist_x.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$10^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_10thhv_3_5um_hist_y.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		random H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_3_5um_hist_y.png}
	\end{minipage}	
	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from simulation results (Distribution Tracking) based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors. The excitation was applied in both planes with an amplitude of 96~nrad. The residual is defined as the final distribution (here $10^4$~turns) minus the initial distribution (here $10^2$~turns), and the ratio as the final distribution divided by the initial distribution. The red line in the residual shows the difference between the Gaussian fit of the distribution and the distribution itself.}
\end{figure}
For the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing the emittance also appears to start at an increased initial value and then stays almost constant. This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution. As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}. Besides an increased emittance, the new distribution also becomes non-Gaussian as can be seen from the residual in respect to the Gaussian distribution in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:patternhist}). For comparison, also the distribution in both planes for a random excitation is shown This initial fast change of the beam distribution was also to the knowledge of the authors for the first time experimentally observed (Sec.~\ref{sec:simex10} and Sec.~\ref{sec:simex7}). The timescale of this fast adjustment is experimentally however much slower than predicted in simulations - on the timescale of minutes compared to seconds.

\subsubsection{10th turn pulsing\label{sec:simex10}}
The $10^{\mathrm{th}}$~turn pulsing was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
\begin{enumerate}
	\item increase of the loss rate with the excitation amplitude (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpsim}),
	\item increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
	\item change of the beam distribution due to the resonant excitation (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpprof})
\end{enumerate}
As loss rates, emittance growth and distribution stayed unchanged for the 4~reference bunches, the above observations can be truly associated with the effect of the resonant excitation.
\begin{figure}[h]
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.49\linewidth]{bunch_intensity_v10th_no_damper_avg_annotate.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{emith_avg_rel_v10th_no_damper_annotate.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{emitv_avg_rel_v10th_no_damper_annotate.png}
	\end{minipage}	
	\caption{\label{fig:10thexp} Relative intensity (top) and emittance (bottom) obtained during the 2016 experiments. For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Sec.~\ref{sec:adt}). The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation as their envelope.}
\end{figure}

@@

Comparing these experimental results with the simulations including only the $10^{\mathrm{th}}$~turn pulsing (solid line, Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth seen experimentally.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit2_rel.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit1.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerrb2uran1_2e-3_10thV_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:10thsim} Relative intensity (top left) and relative vert. emittance (top right), and hor. (bottom left) and vert. (bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$. The solid line indicates an excitation with only $10^{\mathrm{th}}$~turn pulsing in V and the dotted line with $10^{\mathrm{th}}$~turn pulsing in V plus a random dipole noise component in H+V of 6~nrad.}
\end{figure}
Instead, the simulations yield a constant emittance after an initial increase during the first $10^4$~turns or first second. So obviously there is an essential ingredient missing. As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution (dotted line in Fig.~\ref{fig:10thsim}). As a first estimate the value obtained in \cite{md1433_noise_top_energy,md_noise_bbLHC} for 6.5~TeV was used and scaled by the magnetic rigidity to the injection energy of 450~GeV, as no noise estimate is available at injection. This yields a maximum kick amplitude at the transverse damper of approximately
\begin{equation}
\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad},
\end{equation}
Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane only after a first fast adjustment of the beam distribution visible as an increased initial value.

Adding a random noise component has in addition the effect of increasing the loss rate. As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails (Fig.~\ref{fig:10thexpsim}).
\begin{figure}[h]
		\centering
		\includegraphics[width=1.0\linewidth]{scale_amp_10v_ran_10V_sim.png}
	\caption{\label{fig:10thexpsim} Comparison of scaling of loss rates with the $10^{\mathrm{th}}$~turn pulsing excitation amplitude. The experimental results are shown in black (maximum excitation with 48~nrad) and blue (maximum excitation with 96~nrad) and the simulation results (2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad) in red.}
\end{figure}

>>>>

\section{Hollow electron lens for HL-LHC\label{sec:hel}}
\subsection{General Overview\label{sec:hel:intro}}
In general, electron lenses are DC or pulsed low energy electron beams. The electron beam is generated in an electron gun, then guided and confined by strong solenoids and finally dumped onto a collector. Exemplary for the conceptual design of all electron lenses, the HL-LHC HEL is shown in Fig.~\ref{fig:hel_layout}.
\begin{figure}[h]
	\includegraphics[width=1.0\linewidth]{hel_layout_epbeam}% Here is how to import EPS art
	\caption{\label{fig:hel_layout} Layout of HL-LHC HEL.}
\end{figure}
The circulating beam, in case of the LHC the proton beam, is then affected by the electromagnetic field of the electron beam. For the application of active halo control, the electron beam needs to generate an electromagnetic field at the location of the halo particles while leaving the core untouched. This can be achieved by using a uniform hollow distribution in radius $r=\sqrt{x^2+y^2}$ with inner radius $R_1$ and outer radius $R_2$. In this case, the circulating proton beam experiences a radial kick $\theta(r)$
\begin{equation}\label{eq:field_1}
\theta(r)=\frac{f(r)}{(r/R_2)}\cdot \theta_{\rm max},
\end{equation}
where $f(r)$ is a shape function with
\begin{equation}\label{eq:field_2}
f(r) =
\begin{cases} 0 &,\quad r< R_1,\\
\frac{r^2-R_1^2}{R_2^2-R_1^2} &,\quad R_1 \leq r < R_2,\\
1 &,\quad R_2 \leq r
\end{cases}
\end{equation}
and $\theta_{\rm max}=\theta(R_2)$ is the maximum kick angle given by
\begin{equation}
\theta_{\rm max} = \theta(R_2) = \frac{2LI_T(1\pm\beta_e\beta_p)}{4\pi\epsilon_0  \cdot \left(\frac{q}{p_0}\right)_p \cdot \beta_e\beta_p c^2}\cdot\frac{1}{R_2},
\end{equation}
where $L$ is the length of the e-lens, $I_T$ the total electron beam current, $\beta_{e}$ and $\beta_{p}$ the relativistic $\beta$ of electron and proton beam, $\frac{q}{p_0}=\left(B\rho\right)_p$ the magnetic rigidity for the reference particle, $c$ the speed of light and $\epsilon_0$ the vacuum permittivity.
\begin{figure}[b]
	\includegraphics[width=0.8\linewidth]{kick_hel_lhc_no_grid}% Here is how to import EPS art
	\caption{\label{fig:hel_field} Illustration of the hollow electron beam distribution (blue), the kick experienced by the proton beam (red) and the collimators (gray).}
\end{figure}
\begin{table}[t]
	\caption{\label{tab:hel_param}%
		HL-LHC hollow electron lens parameters as in~\cite{hel_cdr}.
	}
	\begin{ruledtabular}
		\begin{tabular}{lcc}
			Geometry & Value& Unit\\
			\colrule
		Length $L$    &  3 & m\\
		Desired range of scraping positions\footnote{$\sigma$ denotes here the Beam sigma and not the collimation $\sigma$ which assumes a normalized emittance of 3.5~$\mu$m.} & 4-8 &$\sigma_p$\\
		\colrule
		Magnetic fields & & \\
		\colrule
		Gun solenoid, $B_g$ & 0.2-0.4 & T\\
		Main solenoid, $B_m$ & 2-6 & T\\
		Collector solenoid, $B_c$ & 0.2-0.4 & T\\
		Compression factor ($k=\sqrt{B_m/B_g}$) & 2.2-5.5 & -\\
		\colrule
		Electron gun & & \\
		\colrule
		Peak yield $I_e$ at 10~keV & 5.0 & A\\
		Gun perveance $P$ & 5 & $\mu$perv\\
		Inner/outer cathode radius, $R_1/R_2$ & 6.75/12.7 & mm\\
		\colrule
		High voltage modulator & & \\
		\colrule
		Cathode-anode voltage & 10.0 & kV\\
		Rise time (10\%-90\%) & 200 & ns \\
		Repetition rate & 35 & kHz
		\end{tabular}
	\end{ruledtabular}
\end{table}

@@

\subsubsection{Prediction of the most efficient pulsing pattern\label{sec:pattern}}
\begin{figure}[b]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_sigm.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit1.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_pattern_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:patternsim} Relative intensity (top left), bunch length (top right) and horizontal (bottom left) and vertical (bottom right) emittance obtained from simulation results (Distribution Tracking) for different pulsing patterns based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors. The excitation was applied in both planes with an amplitude of 96~nrad.}
\end{figure}
In order to predict the most efficient pulsing pattern, the loss rate and emittance growth from the Distribution Tracking are compared for the different pulsing patterns. The results for the 2016 optics and tune are shown in Fig.~\ref{fig:patternsim}. The largest losses are observed for $3^{\mathrm{rd}}$, $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing and the random excitation. As the bunch length decreases accordingly, the losses are mostly associated to off-momentum particles. Only for the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing as well as random excitation a considerable emittance growth is visible, where the random excitation has by far the strongest effect compared to any $k^{\mathrm{th}}$~turn pulsing pattern.

@@

\begin{figure}[t]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$7^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t7skhv_3_5um_hist_x.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		$10^{\mathrm{th}}$ H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_t10skhv_3_5um_hist_y.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		random H+V
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_ranhv_3_5um_hist_x.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		no excitation
		\includegraphics[width=1.0\linewidth]{2016injerra2b2u_3_5um_hist_y.png}
	\end{minipage}	
	\caption{\label{fig:patternhist} Horizontal beam distribution for $7^{\mathrm{th}}$~turn (top left), vertical beam distribution for $10^{\mathrm{th}}$~turn pulsing (top right) and horizontal (bottom left) and vertical (bottom right) beam distribution for random excitation from simulation results (Distribution Tracking) based on the 2016 injection optics with $(Q_x,Q_y)=(64.28,59.31)$ with standard errors. The excitation was applied in both planes with an amplitude of 96~nrad. The residual is defined as the final distribution (here $10^4$~turns) minus the initial distribution (here $10^2$~turns), and the ratio as the final distribution divided by the initial distribution. The red line in the residual shows the difference between the Gaussian fit of the distribution and the distribution itself.}
\end{figure}
For the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing the emittance also appears to start at an increased initial value and then stays almost constant. This behavior is due to the adjustment of the beam distribution during the first $10^4$ turns to a new equilibrium distribution. As the beam distribution is only dumped every $10^4$ turns, this initial fast adjustment manifests itself in an increased initial emittance value in Fig.~\ref{fig:patternsim}. Besides an increased emittance, the new distribution also becomes non-Gaussian as can be seen from the residual in respect to the Gaussian distribution in case of the $7^{\mathrm{th}}$ and $10^{\mathrm{th}}$~turn pulsing (Fig.~\ref{fig:patternhist}). For comparison, also the distribution in both planes for a random excitation is shown This initial fast change of the beam distribution was also to the knowledge of the authors for the first time experimentally observed (Sec.~\ref{sec:simex10} and Sec.~\ref{sec:simex7}). The timescale of this fast adjustment is experimentally however much slower than predicted in simulations - on the timescale of minutes compared to seconds.

\subsubsection{10th turn pulsing\label{sec:simex10}}
The $10^{\mathrm{th}}$~turn pulsing was tested in 2016 with an excitation in the vertical plane only. In summary, the following experimental results were obtained:
\begin{enumerate}
	\item increase of the loss rate with the excitation amplitude (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpsim}),
	\item increase of the emittance growth with the excitation amplitude in the vertical plane but not in the horizontal plane (Fig.~\ref{fig:10thexp}).
	\item change of the beam distribution due to the resonant excitation (Fig.~\ref{fig:10thexp}, \ref{fig:10thexpprof})
\end{enumerate}
As loss rates, emittance growth and distribution stayed unchanged for the 4~reference bunches, the above observations can be truly associated with the effect of the resonant excitation.
\begin{figure}[h]
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.49\linewidth]{bunch_intensity_v10th_no_damper_avg.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{emith_avg_rel_v10th_no_damper.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{emitv_avg_rel_v10th_no_damper.png}
	\end{minipage}	
	\caption{\label{fig:10thexp} Relative intensity (top) and emittance (bottom) obtained during the 2016 experiments. For each maximum amplitude $A_{\mathrm{max}}=48$~nrad and 96~nrad, the excitation amplitude was linearly increased for each group of 4~bunches (see Sec.~\ref{sec:adt}). The different colored solid lines labeled with $n\Delta A$ indicate the average over the 4~bunches with the same excitation amplitude together with the $1\sigma$ standard deviation as their envelope.}
\end{figure}

@@

Comparing these experimental results with the simulations including only the $10^{\mathrm{th}}$~turn pulsing (solid line, Fig.~\ref{fig:10thsim}), the simulations predict qualitatively correctly an excitation amplitude dependent loss rate, but fail to predict the excitation amplitude dependent, continuous emittance growth seen experimentally.
\begin{figure}[h]
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_intensity.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2_rel.png}
	\end{minipage}
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit1.png}
	\end{minipage}	
	\begin{minipage}[t]{0.49\linewidth}
		\centering
		\includegraphics[width=1.0\linewidth]{2016injerra2b2uran1_2e-3_10thV_3_5um_emit2.png}
	\end{minipage}	
	\caption{\label{fig:10thsim} Relative intensity (top left) and relative vert. emittance (top right), and hor. (bottom left) and vert. (bottom right) emittance obtained from simulations (Distribution Tracking) based on the 2016 injection optics with standard errors and $(Q_x,Q_y)=(64.28,59.31)$. The solid line indicates an excitation with only $10^{\mathrm{th}}$~turn pulsing in V and the dotted line with $10^{\mathrm{th}}$~turn pulsing in V plus a random dipole noise component in H+V of 6~nrad.}
\end{figure}
Instead, the simulations yield a constant emittance after an initial increase during the first $10^4$~turns or first second. So obviously there is an essential ingredient missing. As it turns out, the constant emittance can be changed to an excitation amplitude dependent emittance growth by adding a random dipole noise component with uniform distribution (dotted line in Fig.~\ref{fig:10thsim}). As a first estimate the value obtained in \cite{md1433_noise_top_energy,md_noise_bbLHC} for 6.5~TeV was used and scaled by the magnetic rigidity to the injection energy of 450~GeV, as no noise estimate is available at injection. This yields a maximum kick amplitude at the transverse damper of approximately
\begin{equation}
\theta_{\mathrm{random,ADT,max}}(\mathrm{450~GeV}) = 6~\mathrm{nrad},
\end{equation}
Adding this random noise component, the simulations then yield correctly an excitation amplitude dependent emittance growth in the vertical plane only after a first fast adjustment of the beam distribution visible as an increased initial value.

Adding a random noise component has in addition the effect of increasing the loss rate. As the loss rates are in general underestimated in simulations compared to the experimental results, the discrepancy between simulations and experiment is therefore further reduced, but a large difference still prevails (Fig.~\ref{fig:10thexpsim}).
\begin{figure}[h]
		\centering
		\includegraphics[width=1.0\linewidth]{scale_amp_10v_ran_lbllong_sim.png}
	\caption{\label{fig:10thexpsim} Comparison of scaling of loss rates with the $10^{\mathrm{th}}$~turn pulsing excitation amplitude. The experimental results are shown in black (maximum excitation with 48~nrad) and blue (maximum excitation with 96~nrad) and the simulation results (2016 injection optics, standard errors, $(Q_x,Q_y)=(64.28,59.31)$, $10^{\mathrm{th}}$~turn pulsing in V plus random dipole noise in H+V of 6~nrad) in red.}
\end{figure}
